---
title: "T-test_Documentation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{T-test_Documentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(s3oopdh)
```

## Constructor Function - `myttest`
This project aims to create an S3 object-oriented package to execute statistical Student's t-test of two samples, print the output, and then display the distribution on to a boxplot. A constructor function called `myttest()` is created to take in data from sample x and y, as well as alpha (degree of significance) and boolean input whether or not the samples are paired. x and y are constricted to be numeric, alpha to be a double, and paired to be logical input.

According to paired information, the code runs conditional if statement to execute t-test. If `paired==TRUE`, the function will input `paired=TRUE` as the parameter for `t.test`. If `paired==FALSE`, it will run `var.test` to check if the variances are considerably equal or not. Variance equality is assessed through comparison between its p-value and alpha. 

`var.test` has the following null hypothesis:
$$H_0 : \sigma_2^2 / \sigma_1^2 = 1$$
If p-value is less than $\alpha$, variance is considered not equal ($H_0$ rejected) and it will run Welch test. If p-value is equal or greater than $\alpha$, it is considered equal ($H_0$ accepted) and will run two-sample t-test.

t-test runs with the following null hypothesis:
$$H_0 : \mu_2 - \mu_1 = 0$$
Similar to `var.test`, $H_0$ is rejected if p-value < $\alpha$, and vice versa.

Outcomes of t-test are stored in result and each components are stored as list under `obj`, which is then stored under class `Rttest` to be easily retrieved by the subsequent print and plot methods.

```{r}
myttest = function(x = numeric(), y = numeric(), alpha = double(), paired = logical())
{
  if (paired == TRUE)
  {
    result = t.test(x, y, paired=TRUE, conf.level=1-alpha)
  }

  else if (paired == FALSE)
  {
    v = var.test(x,y)

    if (v$p.value < alpha)
    {
      result = t.test(x, y, paired=FALSE, var.equal=FALSE, conf.level=1-alpha)
    }
    else if (v$p.value >= alpha )
    {
      result = t.test(x, y, paired=FALSE, var.equal=TRUE, conf.level=1-alpha)
    }
  }

  #extractable results from t.test

  ci = result$conf.int
  test_type = result$method

  t_value = result$statistic
  df = result$parameter
  p_value = result$p.value
  est_meanordiff = result$estimate
  nullhyp = result$null.value
  stderr = result$stderr
  althyp = result$alternative

  #storing these info into obj of class Rttest

  obj = list(ci = ci,
             test_type = test_type,
             x = x,
             y = y,
             alpha = alpha,
             paired = paired,
             t_value = t_value,
             df = df,
             p_value = p_value,
             est_meanordiff = est_meanordiff,
             nullhyp = nullhyp,
             stderr = stderr,
             althyp = althyp)

  class(obj) = "Rttest"
  obj
}
```

## Method - `print`
This method can easily print useful information from the executed t-test - namely confidence interval, test type, p-value, and the decision whether to reject $H_0$ or not based on its p-value. Other information that has been stored in `Rttest` class can also be retrieved and added to this method to be printed to the console. 

```{r}
print.Rttest = function(x, ...)
{
  ci = x$ci
  test_type = x$test_type
  p_value = x$p_value
  alpha = x$alpha

  if (p_value < alpha)
  {
    RejectNullHypothesis = "Y"
  }

  else
  {
    RejectNullHypothesis = "N"
  }

  output = list(ci = ci, test_type = test_type, pvalue = p_value, RejectNullHypothesis = RejectNullHypothesis)
  output
}
```

## Method - `plot`
This method retrieves the dataset of sample x and y to create boxplot of their distributions. For non-paired sample, this will create two boxplots of corresponding population. For paired sample, this will generate a boxplot of difference between sample x and y. 

``` {r}
plot.Rttest = function(x, ...)
{
  library(ggplot2)
  sample1 = x$x
  sample2 = x$y
  paired = x$paired
  difference = sample1-sample2

  # plot the boxplot of two population samples
  if (paired == FALSE)
  {
    pop = rep(c("sample1", "sample2"))
    distribution = c(sample1, sample2)
    df = data.frame(pop = pop, distribution = distribution)

    g = ggplot(df, aes(x = pop, y = distribution, fill = pop )) + geom_boxplot()
    g = g + ggtitle("Boxplot of Two Population Samples")
    g
  }

  # plot the boxplot of differences of x and y
  else if (paired == TRUE)
  {
    pop = rep(c("difference"))
    df = data.frame(pop = pop, difference = c(difference))

    g = ggplot(df, aes(x = pop, y = difference, fill = pop )) + geom_boxplot()
    g = g + ggtitle("Boxplot of Differences between Two Samples")
    g
  }

}
```

## Example Runs

```{r}
# Run 1
set.seed(32); x=rnorm(30,mean=10,sd=15)
set.seed(35); y=rnorm(30,mean=8,sd=15)
ans1 = s3oopdh::myttest(x, y, alpha = 0.05, paired=FALSE)
print(ans1)
plot(ans1)
# Run 2
set.seed(32); x=rnorm(30,mean=10,sd=5)
set.seed(35); y=rnorm(30,mean=8,sd=15)
ans2 = s3oopdh::myttest(x,y,alpha=0.05,paired=FALSE)
print(ans2)
plot(ans2)
# Run 3
set.seed(32); x=rnorm(30,mean=10,sd=15)
set.seed(35); y = x+ rnorm(30, 5 ,4)
ans3 = s3oopdh::myttest(x,y,alpha=0.05,paired=TRUE)
print(ans3)
plot(ans3)
```
